On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-09-09 LME-MPM.tex
Cambios I (9-9-2020)

\usepackage{lineno}
\usepackage{hyperref}
\usepackage{float}
\usepackage{textcomp}
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}

@@

@@

\newcommand{\CORRECTIONS}[1]{
  \textcolor{red}{{#1}}
}

@@

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under the large
  deformation regime. However, the simulation of shock waves
  propagation in the large deformation regime is still challenging
  under this approach due the incapability of the standard \acrshort{mpm} time
  integration scheme to filter spurious noises. To overcome it, we propose in this paper an explicit Predictor-Corrector time
    integration scheme. Its powerful performance mitigates the
  presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in
  \acrshort{mpm} occurs when the material points cross computational grid boundaries, being caused by
  the lack of smoothness of the interpolation functions. This noise
  results in spurious local variations at the material points, where
  strain-stress fields are computed. This could lead to inaccurate
  solutions as well as aborted simulations in the worst cases. To
  surmount it, we propose in this work the \acrfull{lme} approximation
  schemes as a robust substitute of the traditional shape function in
  \acrshort{mpm}. \acrshort{lme} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter,
  they are able to behave finite element like or as a meshfree
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements through numerical examples.    
\end{abstract}

@@

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}, its popularity
has increased due to its ability to deal with large strain regime
without suffering mesh distortion inaccuracies. One of the main
  fields where this method is strong is the solid dynamics since the original time integration scheme proposed was the \acrfull{fe}~\cite{Sulsky1994}, carried out in an explicit manner. However, in this type or problems, the main instabilities of the original \acrshort{mpm} are even marked.

On the one hand, the main source of instability occurs when
material points cross cell boundaries. This provoked the development
of other interpolation techniques to overcome this limitation such as
the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrated to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is a voxel-based
discretisation technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed by Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the \acrfull{ddmp} method proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases. Therefore \acrshort{ddmp} a large
number of particles is needed for convergence~\cite{DHAKAL2016301}, what makes the method very expensive. \MODIFIED{To avoid the tensile instabilities quite commons in extensions, Sadeghirad {\it et al.} (2011) \cite{Sadeghirad_et_al_IJNME_2011} developed the \acrshort{cpdi} \cite{Sadeghirad_et_al_IJNME_2011}}. In recent years the employment of spline-lines as shape functions has gain popularity with the introduction of the B-Spline 
\acrshort{mpm} proposed by Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employment of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to the \acrshort{mpm} framework with the \acrfull{ctls}
reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}. Unfortunately, when particles are spread
in some particular special patterns, the quality of the \acrshort{ctls} approximation decreases locally.

This document adopts the \acrfull{lme}, or Local \textit{Max-Ent} approximates, as a robust substitute of the aforementioned shape functions in \acrshort{mpm}. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
\acrshort{fem} and meshfree interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The \acrshort{lme} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. \acrshort{lme} basis functions possess many desirable properties for
meshfree algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for related functions~\cite{Arroyo2006}. Important contributions on the Maximum-Entropy have been made by Sukumar and coworkers~\cite{Sukumar15} with Cell-based techniques and the ones carried within the \acrfull{otm} method. The latter methodology has been proven to have a good performance under
the dynamic regime by other researchers, being important the contributions of Li {\it et al.} (2012)~\cite{Li2012} and Navas {\it et al.}
(2018)~\cite{Navas:17b,Navas2018a} in the explicit regime and Navas
{\it et al.}~\cite{Navas2016,Navas2016b,Navas:17c} and Wriggers and
coworkers~\cite{Wriggers18} with implicit schemes. More
recently, under \acrshort{mpm} framework, the work made by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020}. The proposed research delves into the benefits of
the regularization parameter, \gls{LME-beta}, and the analogy of the different obtained shape functions derive by the tuning of this parameter and the traditional \acrshort{mpm} ones.

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretisation technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing somehow the accuracy. One of the first attempts was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} mitigated these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. In a recent publication,
Kularathna \& Soga \cite{Soga_2017} studied an implicit treatment of the pressure in MPM algorithm to simulate material incompressibility
avoiding artificial pressure oscillations by applying
Chorin\textquotesingle s
projection. Within the explicit time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort. In this paper a less time consuming and high efficient explicit predictor-corrector integration method is
proposed. It consists of an accommodation of the traditional \acrfull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.  
\MODIFIED{ Other well known source of energy dissipation and numerical noise is the presence of a non-trivial null space of the linear operator that maps particles values onto nodal values. It was earlier identified by Brackbill~\cite{BRACKBILL1988469} as ringing instability in the \acrfull{pic} method. A recent development in the field of particle methods is a null-space filter engineered by Gritton \&
Berzins~\cite{Gritton2017} which overcome this limitation but still introduce unwelcome damping.  Later, Hammerquist \& Nairn~\cite{HAMMERQUIST2017724} introduced the XPIC as a parametric extension of the aforementioned research with excellent results.}

@@

The article is organised as follows. Section \ref{sec:meshfree-methodology}
is devoted to describe the meshfree methodology adopted in this
research, first \acrshort{mpm} procedure is introduced in
\ref{sec:derivation-mpm}, second the explicit predictor-corrector
time integration scheme is presented in \ref{sec:epc-algor-mpm}, and
third \acrshort{lme} approximation scheme will be introduced in
\ref{sec:local-max-ent}. In Section
\ref{sec:Application-linear-elasticity-dynamic-problems} applications
to prove the numerical accuracy of the proposed approach are
presented. Finally, conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method consists of three main steps: (i) a
variational recovery process, where particle data is projected onto the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention. Three kind of subscripts or superscripts are used
within paper. The subscript $\square_p$ is used to define a particle
variable. While the subscript $\square_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\square^{\psi}$ involves a
virtual magnitude. For the operators, the convention is: $\dot{\square}$ and
$\ddot{\square}$ for the first and second time derivative, $\square \otimes
\square$ means the dyadic operator, $\square \cdot \square$ and $\square \colon \square$ means the
first and second contraction of a tensor, $\Div{\square}$ denotes the
divergence operator, and finally $\Grad{\square}$ and $\GradS{\square}$
denotes the gradient and its symmetric part. Einstein subscripts
convention is adopted therefore repeated index means addition. Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are describe in subsection \ref{sec:derivation-mpm},\ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.

@@

In \acrshort{mpm} the continuum mechanics approach is considered. So on, we will consider a region \gls{domain} occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \Omega
= \Gamma_d \cup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \Omega$
    of the domain $\Omega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) we will relate
  global to local state by the \textit{compatibility equation}, where the the strain field \gls{strain} is extracted from \gls{u} is defined as:
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vec{u}},
\end{equation}
together with essential or \acrfull{dbc} \gls{dirichlet-boundary}. We will further assume that 
strains are 
infinitesimal, therefore second order terms
in the spatial derivatives can be neglected. From here, the
stress field \gls{stress} will be considered the corresponding
conjugate variable for the strain field, being the one which
satisfies (ii) the \textit{conservation of lineal momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vec{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vec{b}
\end{equation}
together with the natural of \acrfull{nbc} \gls{neumann-boundary}.  An aditional component will be (iii) the constitutive equation as a linear
application from $\mathbb{R}^n$ to $\mathbb{R}^n$, which relate stress and strain increments as,
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
In this research, plane strain Linear Elasticity has been
  considered. Thus the constitutive tensor, \tens{D}, is the well
  known linear elastic one. The final equation of the set is (iv) the mass
conservation, which can be obtained by setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vec{v}} = 0.
\end{equation}

In order to obtain the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vec{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vec{u}^{\psi} \in
  \mathcal{H}^1 \mid \vec{u}^{\psi} = \vec{0}\ \text{on}\ \Gamma_d \};
\end{equation}
in which the Cauchy sequences are convergent in \gls{domain} as well:
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vec{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty.
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vec{u} \in
\mathcal{H}^1_0$ such that, for any $\vec{u}^{\psi} \in
\mathcal{H}^1_0$,
the following holds:
%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}}.
\end{equation}\\
%%%%%%%%%%%%%%%%%%%%%
Thus, equation~\eqref{eq:BalanceMomentum_wf}, together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative}, represents the weak form
formulation of the problem. Next, we will discretise the set of equations of the mathematical model using a double discretisation procedure, see figure~\ref{fig:MPM-discretization}.\\

>>>>

\usepackage{lineno}
\usepackage{hyperref}
\usepackage{float}
\usepackage{textcomp}
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}
\usepackage{ulem}
%\usepackage{adjustbox}
\usepackage{rotating}
\usepackage{diagbox}

@@

\newglossaryentry{LME-gamma}
{
  name=$\widehat{\gamma}$,
  description={Dimensionless parameter to control the value of the
    thermalization parameter $\beta$ of the \acrshort{lme} shape functions.}
}

\newglossaryentry{gamma-npc}
{
  name=$\gamma$,
  description={Time integration parameter for the \acrshort{npc} algorithm}
}

@@

\newcommand{\CORRECTIONS}[1]{
  \textcolor{green}{{#1}}
}

@@

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under the large
  deformation regime. However, the simulation of shock waves
  propagation in the large deformation regime is still challenging
  under this approach due \textcolor{green}{to} the incapability of the standard \acrshort{mpm} time
  integration scheme to filter spurious noises. To overcome it, we propose in this paper an explicit Predictor-Corrector time
    integration scheme. Its powerful performance mitigates the
  presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in
  \acrshort{mpm} occurs when the material points cross computational grid boundaries, being caused by
  the lack of smoothness of the interpolation functions. This noise
  results in spurious local variations at the material points, where
  strain-stress fields are computed. This could lead to inaccurate
  solutions as well as aborted simulations in the worst cases. To
  surmount it, we propose in this work the \acrfull{lme} approximation
  schemes as a robust substitute of the traditional shape function in
  \acrshort{mpm}. \acrshort{lme} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter,
  they are able to behave finite element like or as a meshfree
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements through numerical examples.    
\end{abstract}

@@

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}, its popularity
has increased due to its ability to deal with large strain regime
without suffering mesh distortion inaccuracies. One of the main fields where this method is \textcolor{green}{widely used} is solid dynamics.

On one hand, the main source of instability occurs when
material points cross cell boundaries. This \textcolor{green}{issue led to} the development
of other interpolation techniques to overcome this limitation such as
the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrated to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is a voxel-based
discretisation technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed by Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element method due to element distortion.
A more robust alternative is the \acrfull{ddmp} method proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases. Therefore \acrshort{ddmp} \textcolor{green}{requires} a large number of particles for convergence~\cite{DHAKAL2016301}, what makes the method \textcolor{green}{computationally} expensive. \textcolor{blue}{To avoid tensile instabilities that are quite common in extension, Sadeghirad {\it et al.} (2011) \cite{Sadeghirad_et_al_IJNME_2011} developed the \acrshort{cpdi} \cite{Sadeghirad_et_al_IJNME_2011}}. In recent years, the employment of spline-lines as shape functions has \textcolor{green}{gained} popularity with the introduction of the B-Spline 
\acrshort{mpm} proposed by Tielen {\it et al.} (2017)~\cite{TIELEN2017265}, \textcolor{green}{in which} unstructured set of \textcolor{green}{nodes} and
particles \textcolor{green}{can be considered} . More recently, approximants derived from minimization \textcolor{green}{have}
been introduced \textcolor{green}{into} the \acrshort{mpm} framework with the \acrfull{ctls}
reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}. Unfortunately, when particles are spread
in some particular special patterns, the quality of the \acrshort{ctls} approximation decreases locally.

This document adopts the \acrfull{lme}, or Local \textit{Max-Ent} approximates, as a robust substitute of the aforementioned shape functions in \acrshort{mpm}. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
\acrshort{fem} and meshfree interpolations. The \textcolor{green}{\acrshort{lme}}
approximation is based on a compromise between minimizing the
width of the shape function support and maximizing the
entropy of the approximation \textcolor{green}{\cite{Arroyo2006}}. The \acrshort{lme} approximation
may be regarded as a regularization, or \textcolor{green}{(in analogy to statistical mechanics)} \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack \textcolor{green}{or}  uniqueness \textcolor{green}{of}  the triangulation. \acrshort{lme} basis functions possess many desirable properties for
meshfree algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation \textcolor{green}{of affine}  functions~\cite{Arroyo2006}. Important contributions on the Maximum-Entropy have been made by Sukumar and coworkers~\cite{Sukumar15} with Cell-based techniques and the ones carried within the \acrfull{otm} method. The latter methodology has been proven to have a good performance under
the dynamic regime being \textcolor{green}{worth mentioning} the contributions of Li {\it et al.} (2012)~\cite{Li2012} and Navas {\it et al.}
(2018)~\cite{Navas:17b,Navas2018a} in the explicit regime and Navas
{\it et al.}~\cite{Navas2016,Navas2016b,Navas:17c} and Wriggers and
coworkers~\cite{Wriggers18} with implicit schemes. More
recently, under \acrshort{mpm} framework, the work made by Wobbes {\it et al.}(2020)~\cite{Wobbes2020}. The proposed research delves into the benefits of the regularization parameter, \gls{LME-beta}, and the analogy of the different shape functions \textcolor{green}{derived} by the tuning of this parameter and the traditional \acrshort{mpm} ones.

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves, spurious
numerical noises \textcolor{green}{might} appear despite of the employment of these techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretisation technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing somehow the accuracy. One of the first attempts was the proposal of an implicit time integration scheme by Guilkey \& Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} mitigated these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. In a recent publication,
Kularathna \& Soga \cite{Soga_2017} studied an implicit treatment of the pressure in MPM algorithm to simulate material incompressibility
avoiding artificial pressure oscillations by applying
Chorin\textquotesingle s
projection. Within the explicit time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort. In this paper a less time consuming and high efficient explicit predictor-corrector integration method is
proposed. It consists of an accommodation of the traditional \acrfull{npc} scheme, widely employed in Finite Element methods. This method
has \textcolor{green}{already} been chosen \textcolor{green}{in ~\cite{Navas2018a}}  among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and good
performance dealing with solid dynamic problems under a meshfree
framework.  
\textcolor{blue}{Other well known source of energy dissipation and numerical noise is the presence of a non-trivial null space of the linear operator that maps particles values onto nodal values. It was earlier identified by Brackbill~\cite{BRACKBILL1988469} as ringing instability in the \acrfull{pic} method. A recent development in the field of particle methods is a null-space filter engineered by Gritton \&
Berzins~\cite{Gritton2017} which overcome this limitation but still introduce unwelcome damping.  Later, Hammerquist \& Nairn~\cite{HAMMERQUIST2017724} introduced the XPIC as a parametric extension of the aforementioned research with excellent results.}

@@

The article is organised as follows : Section \ref{sec:meshfree-methodology}
is devoted to describe the meshfree methodology adopted in this
research, first \acrshort{mpm} procedure is introduced in
\ref{sec:derivation-mpm}, second the explicit predictor-corrector
time integration scheme is presented in \ref{sec:epc-algor-mpm}, and
third \acrshort{lme} approximation scheme will be introduced in
\ref{sec:local-max-ent}. In Section
\ref{sec:Application-linear-elasticity-dynamic-problems}\textcolor{green}{,} applications to prove the numerical accuracy of the proposed approach are
presented. Finally, conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method consists of three main steps: (i) a
variational recovery process, where particle data is projected onto the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention: \textcolor{blue}{Any vectorial field is defined with an overarrow as $\vec{\square}$}. Three kind of subscripts or superscripts are used within paper. The subscript $\square_p$ is used to define a particle
variable. While the subscript $\square_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\square^{\psi}$ involves a
virtual magnitude. For the operators, the convention is: $\dot{\square}$ and
$\ddot{\square}$ for the first and second time derivative, $\square \otimes
\square$ means the dyadic operator, $\square \cdot \square$ and $\square \colon \square$ means the
first and second contraction of a tensor, $\Div{\square}$ denotes the
divergence operator, and finally $\Grad{\square}$ and $\GradS{\square}$ denotes the gradient and its symmetric part. Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are \textcolor{green}{described} in subsection \ref{sec:derivation-mpm}, \ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.

@@

In \acrshort{mpm} the continuum mechanics approach is considered. We consider a region \gls{domain} occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \Omega
= \Gamma_d \cup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \Omega$
    of the domain $\Omega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) we relate global to local state by the \textit{compatibility equation} \textcolor{green}{where the strain field \gls{strain} is defined by}:
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vec{u}},
\end{equation}
together with essential or \acrfull{dbc} \textcolor{green}{at} \gls{dirichlet-boundary}. We will further assume that 
strains are infinitesimal. The stress field \gls{stress} is considered the corresponding
conjugate variable for the strain field, being the one which
satisfies (ii) the \textit{conservation of \textcolor{green}{linear} momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vec{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vec{b}
\end{equation}
together with the natural of \acrfull{nbc} \textcolor{green}{at} \gls{neumann-boundary}. An \textcolor{green}{additional} component \textcolor{green}{is} (iii) the constitutive equation which relates stress and strain increments as,
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\textcolor{green}{\Delta} \sigma} = \tens{D} \colon \tens{\textcolor{green}{\Delta} \varepsilon}.
\end{equation}
In this research, plane strain linear elasticity \textcolor{green}{is} considered. The final equation of the set is (iv) the mass conservation, which can be \textcolor{green}{expressed by} 
\begin{equation}
  \label{eq:Rho-material-derivative}
    \dot{\rho} + \rho \Div{\vec{v}} = 0.
\end{equation}

In order to obtain the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vec{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vec{u}^{\psi} \in
  \mathcal{H}^1 \mid \vec{u}^{\psi} = \vec{0}\ \text{on}\ \Gamma_d \};
\end{equation}
\textcolor{green}{where $\vec{u}^{\psi} \in \mathcal{H}^1$ means that}
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vec{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty.
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vec{u} \in
\mathcal{H}^1_0$ such that, for any $\vec{u}^{\psi} \in
\mathcal{H}^1_0$,
the following holds:
%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%
Thus, equation~\eqref{eq:BalanceMomentum_wf}, together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative}, represents the weak form
formulation of the problem. Next, we will discretise the set of equations of the mathematical model using a double discretisation procedure, see figure~\ref{fig:MPM-discretization}.\\

\textcolor{blue}{First, the velocity field and the virtual displacements fields are discretised in a finite set of nodes $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^{2}$, where $\mathcal{B} = 1, \ldots, n_I$. An approximation of the continuum field can be obtained with the help of nodal values and an appropriate interpolation functions $N_I(\vec{x})$. Also spatial derivatives of those quantities, such as gradients and divergences, are computed through the support of the background set of nodes as}
    \label{eq:variable_reconstruction}
    \varphi(\vec{x}) &= \sum_{I\ \in\ \mathcal{B}} N_I(\vec{x})\ \varphi_I \\
    \label{eq:grad_variable_reconstruction}
    \Grad{\varphi}(\vec{x}) &= \sum_{I\ \in\ \mathcal{B}} \varphi_I\ \otimes \Grad{N_I(\vec{x})}
\textcolor{blue}{Secondly, the continuum \gls{domain} is discretised with a finite set of material points (also denominated particles the manuscript) $\hat{\Omega} = \{ \vec{x}_p, p \in \mathcal{C} \} \subset \Omega$, where $\mathcal{C} = 1, \ldots, n_p$. Any particle field such as position, velocity, mass, volume and stress denoted by $\vec{x}_p$, $\vec{v}_p$, $m_p$, $V_p$ and $\tens{\sigma}_p$, respectively, are assigned to each material point. Furthermore, any particle field $\varphi_p$ can be approximated within the nodes in the neighbourhood of each particle, $\mathcal{B}_p \subset \mathcal{B}$, and evaluating the interpolation function in the position of each particle as}
    \label{eq:particle_variable_reconstruction}
\varphi_p &= \sum_{I\ \in\ \mathcal{B}_p} N_I(\vec{x}_p)\ \varphi_I \\
\Grad{\varphi_p} &= \sum_{I\ \in\ \mathcal{B}_p} \varphi_I \otimes \Grad{N_I(\vec{x}_p)}
\textcolor{blue}{The integrals appearing in the weak form \eqref{eq:BalanceMomentum_wf} are evaluated by means of the Riemann integral \cite{Riemann_1854} applied to a finite set of points with associated volumes $V_p$ interpreted as quadrature weights.}
    \label{eq:particle_riemann_integral}
\Integral{3}{\varphi} = \sum_{p\ \in\ \mathcal{C}} \varphi_p\ V_p 
